---
title: "cARE Stats"
output: html_notebook
---

For CARE dataset, mean of engagement
```{r}
mean(care_data$engagement)
```

For PNS dataset, mean of engagement
```{r}
mean(pns_data$engagement)
```

For CARE dataset, mean of current ema engagement

```{r}
aggregate(engagement ~ previous_ema_engagement, data = care_data, mean)
```

For PNS dataset, mean of current ema engagement for previous ema engagement

```{r}
aggregate(engagement ~ previous_ema_engagement, data = pns_data, mean)
```

`
For CARE dataset, mean of EMA count per participant

```{r}
care_data_ema_count <- aggregate(Initiated ~ Part_ID, data = care_data, length)
mean(care_data_ema_count$Initiated)
```

```{r}
median(care_data_ema_count$Initiated)
```
```{r}
sd(care_data_ema_count$Initiated)
```

```{r}
min(care_data_ema_count$Initiated)
```

```{r}
max(care_data_ema_count$Initiated)
```

```{r}
IQR(care_data_ema_count$Initiated)
```


For PNS dataset, mean of EMA count per participant

```{r}
pns_data_ema_count <- aggregate(delivered_hrts ~ id, data = pns_data, length)
mean(pns_data_ema_count$delivered_hrts)
```
```{r}
median(pns_data_ema_count$delivered_hrts)
```

```{r}
sd(pns_data_ema_count$delivered_hrts)
```

```{r}
min(pns_data_ema_count$delivered_hrts)
```

```{r}
max(pns_data_ema_count$delivered_hrts)
```

```{r}
IQR(pns_data_ema_count$delivered_hrts)
```

For CARE dataset, summary statistics table

```{r}
care_data$new_day_modified <- care_data$new_day + 7
```

```{r}
engagement_rate_per_day_df = aggregate(engagement ~ new_day_modified, data = care_data, mean)
colnames(engagement_rate_per_day_df)[2] <- "mean_engagement"
engagement_rate_per_day_df
```

```{r}
atleast_one_ema_df = aggregate(Part_ID ~ new_day_modified, data = care_data, n_distinct)
colnames(atleast_one_ema_df)[2] <- "total_individuals_with_atleast_one_ema"
atleast_one_ema_df
```

```{r}
total_ema_per_day_df  = aggregate(Initiated ~ new_day_modified, data = care_data, length)
colnames(total_ema_per_day_df)[2] <- "total_ema"
total_ema_per_day_df
```

```{r}
care_comb_df <- cbind(engagement_rate_per_day_df,atleast_one_ema_df,total_ema_per_day_df)
care_comb_df <- care_comb_df %>% select(-3, -5) 
colnames(care_comb_df)[1] <- "day"
care_comb_df
```

```{r}
library(readr)
write_csv(care_comb_df, "O:/QMP-BIDS/Utah - CARE Datasets/Aditi/CARE/Code/Survival-HMM-Care/summary_stats_care_pns/care_summary_stats.csv")
```


For PNS dataset, summary statistics table

```{r}
pns_data$study_day_modified <- pns_data$study_day + 7
```

```{r}
engagement_rate_per_day_df = aggregate(engagement ~ study_day_modified, data = pns_data, mean)
colnames(engagement_rate_per_day_df)[2] <- "mean_engagement"
engagement_rate_per_day_df
```

```{r}
atleast_one_ema_df = aggregate(id ~ study_day_modified, data = pns_data, n_distinct)
colnames(atleast_one_ema_df)[2] <- "total_individuals_with_atleast_one_ema"
atleast_one_ema_df
```

```{r}
total_ema_per_day_df  = aggregate(record_id ~ study_day_modified, data = pns_data, length)
colnames(total_ema_per_day_df)[2] <- "total_ema"
total_ema_per_day_df
```

```{r}
pns_comb_df <- cbind(engagement_rate_per_day_df,atleast_one_ema_df,total_ema_per_day_df)
pns_comb_df <- pns_comb_df %>% select(-3, -5) 
colnames(pns_comb_df)[1] <- "day"
pns_comb_df
```

```{r}
library(readr)
write_csv(pns_comb_df, "O:/QMP-BIDS/Utah - CARE Datasets/Aditi/CARE/Code/Survival-HMM-Care/summary_stats_care_pns/pns_summary_stats.csv")
```



```{r}
care_data_individual <- aggregate(data = care_data, engagement ~ Part_ID, mean)
ggplot(data = care_data_individual, aes(x = engagement)) + geom_histogram()
```

```{r}
mean(care_data_individual$engagement)
```

```{r}
sd(care_data_individual$engagement)
```

```{r}
IQR(care_data_individual$engagement)
```
```{r}
pns_data_individual <- aggregate(data = pns_data, engagement ~ id, mean)
ggplot(data = pns_data_individual, aes(x = engagement)) + geom_histogram()
```

```{r}
mean(pns_data_individual$engagement)
```

```{r}
sd(pns_data_individual$engagement)
```

```{r}
IQR(pns_data_individual$engagement)
```
Method 1: To plot histograms of CARE and PNS on the same axis

```{r}
c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

b <- min(c(care_data_individual$engagement,pns_data_individual$engagement)) # Set the minimum for the breakpoints
e <- max(c(care_data_individual$engagement,pns_data_individual$engagement)) # Set the maximum for the breakpoints
ax <- pretty(b:e, n = 12) # Make a neat vector for the breakpoints
ax
```

```{r}
hgA <- hist(care_data_individual$engagement, breaks = ax, plot = FALSE) # Save first histogram data
hgB <- hist(pns_data_individual$engagement, breaks = ax, plot = FALSE) # Save 2nd histogram data

plot(hgA, col = c1) # Plot 1st histogram using a transparent color
plot(hgB, col = c2, add = TRUE) # Add 2nd histogram using different color
```

Method 2: To plot histograms of CARE and PNS on the same axis

```{r}
care_data_individual <- aggregate(data = care_data, engagement ~ Part_ID, mean)
pns_data_individual <- aggregate(data = pns_data, engagement ~ id, mean)

pns_data_individual <- pns_data_individual %>% rename(Part_ID = id)

care_data_individual$source="CARE"
pns_data_individual$source="PNS"

dat_combined = rbind(care_data_individual, pns_data_individual)
dat_combined

```

```{r}
ggplot(dat_combined, aes(engagement, fill=source)) + 
  geom_histogram(aes(y=..density..), position="identity", colour="grey40", alpha=0.5, binwidth=0.01) +
  ggtitle("Histograms of individual response rates across all days") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Response Rate") +
  ylab("Density")
```
```{r}
care_data_individual <- aggregate(data = care_data, Initiated ~ day_time_delta, length)
pns_data_individual <- aggregate(data = pns_data, delivered_hrts ~ day_time_delta, length)

care_data_individual <- care_data_individual %>% rename(Frequency = Initiated)
pns_data_individual <- pns_data_individual %>% rename(Frequency = delivered_hrts)

care_data_individual$source="CARE"
pns_data_individual$source="PNS"

dat_combined = rbind(care_data_individual, pns_data_individual)
dat_combined
```

```{r}
care_data_individual
```

```{r}
pns_data_individual
```


```{r}
max(pns_data_individual$Frequency)
```

```{r}
aggregate(data = pns_data, id ~ day_time_delta, length)
```
```{r}
pns_data %>% group_by(day_time_delta) %>% summarize(n())
```

```{r}
ggplot(dat_combined, aes(day_time_delta, fill=source)) + 
  geom_histogram(aes(y=..density..), position="identity", colour="grey40", alpha=0.5, binwidth=0.01) +
  geom_density() +
  ggtitle("Histograms of frequency of different deltas") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Day Time Delta") +
  ylab("Density")
```

```{r}
ggplot(dat_combined, aes(day_time_delta, fill=source)) + 
  geom_histogram(position="identity", colour="grey40", alpha=0.5, binwidth=0.01) +
  ggtitle("Histograms of frequency of different deltas") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Day Time Delta") +
  ylab("Frequency")
```
```{r}
hist(pns_data_individual$day_time_delta)
```


```{r}
# Histogram with kernel density
ggplot(care_data, aes(x = day_time_delta)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(lwd = 1, colour = 4,
               fill = 4, alpha = 0.25)
```
```{r}
care_data$source="CARE"
pns_data$source="PNS"

dat_combined = rbind(care_data, pns_data)
dat_combined
```
```{r}
ggplot(dat_combined, aes(day_time_delta, fill=source)) + 
  geom_histogram(position="identity", colour="grey40", alpha=0.5, binwidth=0.01) +
  ggtitle("Histograms of frequency of different deltas") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Day Time Delta") +
  ylab("Frequency")
```

```{r}
# Histogram with kernel density
ggplot(pns_data, aes(x = day_time_delta)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(lwd = 1, colour = 4,
               fill = 4, alpha = 0.25)
```

```{r}
# Histogram with kernel density
ggplot(care_data_individual, aes(x = day_time_delta)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(lwd = 1, colour = 4,
               fill = 4, alpha = 0.25)
```

